Dark matter halo shape at $z=0$ in the Auriga simulations: radial evolution and alignment with the stellar disk

### 2019-07-19 Document/paper.tex
edit

Our Figure \label{fig:observations} shows an explicit comparison in
the $c/a$-$b/a$ plane of our results in MHD simulations against the
results by LM10 and BBFK16.  
We find six MHD halos with $b/a<0.93$ and $c/a>0.77$ that could be
considered consistent with their shape constraints  by BBFK16, while
only one outlier DMO halo is consistent with those constraints.
In contrast, none of the simulated halos (MHD nor DMO) is consistent
with the LM10 results. 
The change of triaxility with radius in our simulations cannot account
for these two extremely different observational constraints at
different radii.

@@

From the results presented in this work we advance the idea that the
twisting shells in the density is a feature that deserves to be
explored in the process of constraining shape parameters from tidal
stream data. 
The current parameterizations of coaxial density/potential shells
might be too rigid.
The inclusion of a new free parameter describing this degree of
twisting might relax the conflict between the observational
constraints and the numerical results.

{{\bf alinenaciones y twisting as a function of environment, 
the difference between accretting and stalled halos, }

%Porciani: stalled halos keep their dark matter density profile
%while accreting halos have. Shells don't pile up in accreting halos, 
%they approximately do in stalled halos}

%Porciani: The stalled halos are the most similar to the MW at least in
%terms of the velocity anisotropy parameter an should be embeded in a
%thick filament of tha large scale structure.

%IDEA PARA YEIMY: medir scatter en stellar to DM halo mass ratio as a
%function of the anisotropy parameter or the persistence parameter in 
%Illustris. Primer proxy para el galaxy formation bias. Esto se puede
%sacar pronto. (ademas otra medida de los compacto que es la distribucion
%estelar)
%esto puede ser comparado con los datos de vipers que tiene un sesgo 
%donde se ven diferentes clustering a misma masa para galaxias mas o
%menos concentradas

%otra idea: calcular la entropia del beta skeleton para distinguir 
%el clustering de dos muestras diferentes. i.e. la entropia para
%caracterizar assembly bias.

@@

In this paper we measured the shape of thirty isolated Milky Way like 
dark matter haloes simulated in the Auriga project using the zoom-in
technique. 
We also used six of these haloes that were simulated at higher
resolution. 
The Auriga project simulated these halos with two different setups:
dark matter only (DMO) simulation and full magnetohydrodynamics (MHD)
includeing star formation and feedback.
We used the shape measurement algorithm by \cite{Allgood06} on the
dark matter halos of these two kinds of simulations.

We find that MHD halos are rounder than DMO halos every sampled radii. 
MHD halos tend towards more oblate shapes (T < 0.5)
despite DM halos tendency towards more prolate shapes (T>0.5). 
It is also noticeable that the rounding effect of baryons is stronger
at the disk regime, where the haloes are close to oblate.

This rounding effect can be explained by gravitational effect of the
flattened axisymmetric galactic disk. 
It also explanes the weakness of this affect around $\approx 100$kpc,
where the disk potential is weaker compared to the DM halo potential.

The second part of our study on the historical shape finds that in the
presence of baryons the memory effect found by \citep{VeraCiro11} does
not hold for all halos in the high resolution sample.

>>>>

Our Figure \label{fig:observations} shows an explicit comparison in
the $c/a$-$b/a$ plane of our results in MHD simulations against the
results by LM10 and BBFK16.  
We find six MHD halos with $b/a<0.93$ and $c/a>0.77$ that could be
considered consistent with their shape constraints  by BBFK16, while
only one outlier DMO halo is consistent with those constraints.
In contrast, none of the simulated halos (MHD nor DMO) is consistent
with the LM10 results. 
The change of triaxility with radius in our simulations cannot account
for these two extremely different shape constraints at different
radii.

@@

From the results presented in this work we advance the idea that the
twisting shells in the density is a feature that deserves to be
explored in the process of constraining shape parameters from tidal
stream data. 
The inclusion of a parameterization describing this degree of
twisting might relax the conflict between the observational
constraints and the numerical results.

From the purely numerical point of view a better understading of the
disk-halo alignment details will certainly require to study how the
halo and the disk co-evolved as a function of time.
The effect of the cosmic web (both dark matter and gaseous)
\citep{2014MNRAS.443.1090F,2017MNRAS.469..594B,2019MNRAS.487.1607G} could also be in
order to gain a full picture.

@@

In this paper we measured the shape of thirty isolated Milky Way like 
dark matter haloes simulated in the Auriga project using the zoom-in
technique. 
The Auriga project simulated these halos with two different setups:
dark matter only (DMO) simulation and full magnetohydrodynamics (MHD)
including star formation and feedback.
We used the shape measurement algorithm by \cite{Allgood06} on the
dark matter halos of these two kinds of simulations to quantify the
halo shape as a function of radius and the degree of aligment between
the stellar disk and the halo.

We find that MHD halos are rounder than DMO halos every sampled radii. 
MHD halos tend towards more oblate shapes ($T< 1/3$) DMO halos
tend towards more prolate shapes ($T>2/3$).  
The rounding effect is more noticeable as one moves closer to galactic
disk and strongly  correlates with baryonic properties of the disk
$\leq 0.10R_{200}\approx 20$kpc. 
More precisely, the triaxility is smaller for large and massive
stellar disks with low gas densities in its core $\leq 1$ kpc.

It also explanes the weakness of this affect around $\approx 100$kpc,
where the disk potential is weaker compared to the DM halo potential.
